Annotating, Understanding, and Predicting Long-term Video Memorability

### 2018-04-27 ARTICLE.tex
Final version of accepted paper after reviewing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{booktabs} % For formal tables
%% perso packages:
\usepackage{epstopdf}
\usepackage{microtype}
\usepackage{tabularx}
\usepackage{multirow}
%\usepackage{arydshln}
\usepackage{subfig}
\usepackage{balance}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT's Identification
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CONFERENCE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Conference
\acmConference[ICMR 2018]{ACM Yokohama conference}{June 2018}{Yokohama, Japan} 
\acmYear{2018}
%\copyrightyear{2018}
%\acmArticle{4}%??
%\acmPrice{15.00}%??

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Annotating, understanding, and predicting long-term video memorability}
%\titlenote{}
%\subtitle{}
%\subtitlenote{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%author 1
%\author{Romain Cohendet}
%\authornote{}
%\orcid{1234-5678-9012}
%\affiliation{%
%	\institution{Technicolor}
%	\streetaddress{Technicolor Research \& Licensing, 975 Avenue des Champs Blancs, 35576 Cesson-Sevigne, France}
%	\city{Rennes} 
%	\state{France} 
%	\postcode{}
%}
%\email{romain.cohendet@technicolor.com}

%author 2
%\author{Karthik Yadati}
%\affiliation{%
%	\institution{University of Delft}
%	\streetaddress{}
%	\city{Delft} 
%	\state{} 
%	\postcode{}
%}
%\email{N.K.Yadati@tudelft.nl}

%author 3
%\author{Ngoc Khanh Duong}
%\affiliation{%
%	\institution{Technicolor}
%	\city{Rennes} 
%	\country{France}}
%\email{Quang-Khanh-Ngoc.Duong@technicolor.com}

%author 4
%\author{Claire-Helene Demarty}
%\affiliation{%
%	\institution{Technicolor}
%	\city{Rennes}
%	\country{France}
%}
%\email{Claire-Helene.Demarty@technicolor.com}

%%%%%%%%%%%%
%%% for blind review %%%
%author 1
%\author{Removed for review}
%author 2
%\author{Removed for review}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{Removed for blind review}%{R. Cohendet et al.}%Removed for blind review

@@

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% KEYWORDS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\keywords{Video memorability, Long-term memory, Measurement protocol, Deep learning, Multimedia information retrieval}

@@

%% About the ICMR article on Media retrieval with VM
	% focus on "Multimedia content-based retrieval"
	% Length: from 6 to 8 pages + 1 additional page for references (no longer distinction between long and short papers)
	% Submission deadline: 2018/01/20
	% Contributions addressing the challenges of large-scale search and user behavior analysis are especially welcome

@@

%objectives of the present study
While participating in the expansion of research on VM, the contributions of this paper are threefold:
\begin{itemize}
	\item We propose a new protocol that measures very long-term memory performances (from weeks to years after memorization) to collect ground truth data of good quality and make the corresponding dataset available for the research community (section \ref{section_dataset_construction}).
	\item We assess and analyze the collected data and provide useful insights on the understanding of VM (section \ref{study_memorability}).
	\item We build a computational model based on machine learning techniques which allows to predict VM score of a given video. For this purpose, we investigate the use of various types of audio and visual features, ranging from low-level characteristics to emotional and (highly) semantic features (section \ref{mem-pred}).
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Previous work}
In this section, we review previous work on image and video memorability prediction, focusing on annotation protocols and modelling aspects.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Measurement of memorability}
%work in psychology
Long-term memory has been studied for over a century in psychology, from multiple perspectives, since the seminal experimental studies of Ebbinghaus \cite{ebbinghaus_1913_memory} until the more recent neuro-imaging studies \cite{blumenfeld_2007_prefrontal}.
This work provided researchers, interested in computational understanding of IM and VM, with several memory tests (see \cite{richardson_1988_measures} for an extensive overview), such as the recognition test \cite{isola_2011_makes,khosla_2015_understanding,han_2015_learning} or the textual questions-based recall survey \cite{shekhar_2017_show}.
It also demonstrated that humans have an extensive long-term visual memory, which enables them to recall a great amount of images \cite{standing_1973_learning} and image details \cite{brady_2008_visual}, as well as of videos \cite{furman_2007_they}.
This is one of the observations at the origin of the work on memorability in computer science \cite{isola_2011_makes}.
Several factors have also been highlighted for their critical influence on long-term memory, such as emotion \cite{kensinger_2008_memory}, semantics \cite{quillan_1966_semantic}, several demographic factors \cite{cohendet_2016_using}, memory re-evocation \cite{nadel_1997_memory}, or passage of time \cite{mcgaugh_2000_memory}.
These different factors are all important to better understand memorability, and are found valuable computational features for image and video memorability prediction.

@@

%Shekhar et al.
In another earlier approach for VM measurement, the participants performed a crowdsourcing experiment that consisted of a free viewing task during which they saw a series of videos, followed by a recall task in which they had to answer textual questions (such as: "Did you see a man juggling?", "Did you see a car on road?") \cite{shekhar_2017_show}.
The major drawback of the study comes from the use of questions instead of a classic visual recognition task.
Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty.
The authors tested the complexity of the chosen textual questions using the Flesch-Kincaid Grade Level readability metric, which is designed to quantify how difficult to understand an English text is \cite{kincaid_1975_derivation}.
However, this measure might not be sufficient to represent the questions in all their complexities, such as the ease for a person to draw images from the words, to connect the words to the associated concepts in memory, to associate the words to the corresponding scene, etc.
Moreover, there is an unequal distribution of the reading comprehension of English among Amazon Mechanical Turk workers, who are not necessary anglophone people.
This problem of a potential unequal complexity of the question becomes even more important in view of the use by authors of the response time to calculate memorability scores, which might also critically depend on the complexity of the questions.
One will note that this potential bias could have also affected the measure of inter-human consistency.
Furthermore, the questions were handcrafted, and the choices for types of questions and videos were very limited (e.g., the question "Did you see a car on road?" implies that in a whole experimental session only one car on a road should appear to connect the memory performance to one particular video).
The authors also manually ensured that no textual questions nor videos in a session were similar in content.
This makes it difficult to generalize this protocol to the construction of an extensive dataset.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memorability score calculation}
%Each video has also been viewed as a filler by $10.5$ participants.
After collecting the data, we kept only the 660 videos that had been seen at least 4 times as targets (from the initial set of 700 videos).
On average, each video of our dataset has been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enters in the calculation of a memorability score.
We then assigned a memorability score to each video, defined as the correct recognition rate of the video when viewed as target.
The average percentage of correct detections for all participants was $46.71\%$ (stdev = 14.65\%), and the average false alarm rate (i.e., the percentage of answers on fillers) was $4.16\%$ (stdev = 5.27\%).
Figure \ref{fig:response_time_vs_memorability}(a) provides a distribution of the videos according to their degree of memorability.
The dataset is publicly released here\footnote{\textit{www.remove-for-blind-review}. The original videos are not released because of copyright issues, but instead we provide the process we used to extract them from the movies and the features used in section \ref{mem-pred}.}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Consistency analysis}
\label{ssec:human}
%global consistency calculation
We follow the method proposed in \cite{isola_2014_makes} to measure human consistency when assessing memorability of videos.
We randomly split our 104 participants into two independent groups of equal size, and calculate how well video memorability scores from the first group of participants match with video memorability scores from the second group.
Averaging over 25 random half-split trials, an average Spearman's rank correlation (i.e., a global human consistency) of $0.57$ is observed between these pairs of scores.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Neutral and typical videos}
In our experiment, participants were given clear instructions that they had to really recognize any video they were presented as already seen, and not only guess that a video was coming from a movie whose title was proposed in the questionnaire.
In this section, we perform an analysis to compare neutral videos, which contain no element that would enable participants to guess that a video belongs to a particular movie, and typical videos.
Indeed, if neutral videos received objective answers from participants, it might be more subjective for typical videos, that could be more or less easily related to the movie they belong to.

A Wilcoxon rank-sum test indicated that the memorability was greater for neutral ($Mdn=.20, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, with $Z=10.22, p<.00001$.
Apart from the subjectivity aspect, we expected such a result because neutral videos contain less contextual elements, useful for recognition.
Thus, this result does not necessarily mean that participants tended to guess -- rather than simply recognize -- videos selected from movies they have seen.

As for the human consistency on memorability, a Wilcoxon rank-sum test indicated that it was slightly greater for neutral ($Mdn=.44, \mu=.45$) than for typical ($Mdn=.40, \mu=.41$) videos, with $Z=2.75, p<.01$.
Along with the comments collected from the participants, who have as a majority reported difficulty to know if they were guessing or really recognizing the videos, this result suggests that human congruency is higher for more 'objectively' recognized segments than for ones with subjectivity as part of the equation. 
One could note there is probably not just pure subjectivity here.
Context might have biased the results as it might have helped some participants in their recognition task.
If confirmed, this would constitute a weakness of our protocol to collect extensive data, that one should in that case counteract by adapted measures of control.

@@

In \cite{shekhar_2017_show}, the response time of the participants was taken to be the measure of video memorability.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{New manner to compute memorability scores: take into account time and FA}
%Une nouvelle manière pour prendre en compte les FA dans les scores de mémorabilité applicable à l'étude de la mémorabilité sur les images
%Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.
% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the movie --> No' --> So the point here is just to provide an overview.

@@

\subsection{Modelling and evaluation metric}
\label{model-eval}
We use the features discussed in Section \ref{feat-extr} to train a Support Vector Regression (SVR) model for the VM score prediction. The choice of SVR is guided by the nature of the problem as well as by the small size of the dataset. We have chosen to go with the same regressor for all the features because our focus is mainly on identifying which features are more important for VM prediction. This way we ensure that the difference in performances is because of the features themselves. Note that, in addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating them into a single feature vector.
While performing such a concatenation, we use the low-dimension version of the features for C3D and SentiBank features, obtained after applying a dimension reduction method (PCA) to the original set. In the experiments where we use PCA for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.

@@

As it can be seen, visual semantic features derived from image captioning clearly offer better prediction results compared to all other considered features. This is not surprising as they capture visual attributes along the scene, which are known to play an important role in human memory. This is also inline with a previous study in IM \cite{squalli_2017_deep} where image captioning-based features helped better predict IM than the conventional CNN features. %These best prediction results (i.e., 0.31 and 0.38 for the test sets) are quite close to the human consistency shown in Figure \ref{fig:human_consistency}b (i.e. about 0.36 when having 4 annotations per video and about 0.47 when having 8 annotations per video).
When predicting VM on the test set, C3D features showed to be quite effective to the task as they encode the visual spatio-temporal information of the video. On the contrary, audio information captured by AudioSet features does not seem to be enough for VM prediction. One of our initial hypotheses was that emotion would play an important role in VM, supported by literature from psychology \cite{emo-mem}.
However, when observing the results in Table \ref{res-4-10-ann}, both Affect \cite{affect} and SentiBank \cite{sb-feat} performed quite poorly compared to the image captioning features.
%Observing the first set of results in Table \ref{res-4-10-ann}, where we have at least 4 annotations per video, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set. 
%Image captions capture the visual semantics in the video, while C3D features encode the visual spatio-temporal information.
%These results are consistent with previous results on predicting image memorability \cite{squalli_2017_deep}.
%Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
%Other features like AudioSet come close to the performance of the C3D, but only audio signal does not seem to be enough for predicting video memorability.
Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features. One of the reasons for this could be that there is a lot of redundance when combining all the features into a single feature vector, or it could be that the size of the combined feature vector is too big when compared to the size of the dataset. Thus, in future work we could look at selectively combining the features to investigate if that improves the performance.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{booktabs} % For formal tables
%% perso packages:
\usepackage{epstopdf}
\usepackage{microtype}
\usepackage{tabularx}
\usepackage{multirow}
%\usepackage{arydshln}
\usepackage{subfig}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{amsmath}
%\usepackage{times}
\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT's Identification
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\copyrightyear{2018} 
\acmYear{2018} 
\setcopyright{acmcopyright}
\acmConference[ICMR '18]{2018 International Conference on Multimedia Retrieval}{June 11--14, 2018}{Yokohama, Japan}
\acmBooktitle{ICMR '18: 2018 International Conference on Multimedia Retrieval, June 11--14, 2018, Yokohama, Japan}
\acmPrice{15.00}
\acmDOI{10.1145/3206025.3206056}
\acmISBN{978-1-4503-5046-4/18/06}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Annotating, Understanding, and Predicting Long-term Video Memorability}
%\titlenote{}
%\subtitle{}
%\subtitlenote{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%author 1
\author{Romain Cohendet$^{\ast}$}{\thanks{$^{\ast}$Both authors contributed equally to this work.}}
%\authornote{}
%\orcid{1234-5678-9012}
\affiliation{%
	\institution{Technicolor}
	\streetaddress{Technicolor Research \& Licensing, 975 Avenue des Champs Blancs, 35576 Cesson-Sevigne, France}
	\city{Rennes}
	\state{France}
	\postcode{}
}
\email{Romain.Cohendet@technicolor.com}

%author 2
\author{Karthik Yadati$^{\ast}$}
	\affiliation{%
	\institution{Delft University of Technology }
	\streetaddress{}
	\city{Delft} 
	\state{} 
	\postcode{}
}
\email{N.K.Yadati@tudelft.nl}

%author 3
\author{Ngoc Q. K. Duong}
\affiliation{%
	\institution{Technicolor}
	\city{Rennes} 
	\country{France}}
\email{Quang-Khanh-Ngoc.Duong@technicolor.com}

%author 4
\author{Claire-H\'el\`ene Demarty}
\affiliation{%
	\institution{Technicolor}
	\city{Rennes}
	\country{France}
}
\email{Claire-Helene.Demarty@technicolor.com}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{R. Cohendet et al.}%Removed for blind review

@@

%\ccsdesc[500]{Computing methodologies~Computer vision}
%\ccsdesc[500]{Computing methodologies~Scene understanding}
%\ccsdesc[300]{Computing methodologies~Visual content-based indexing and retrieval}
%\ccsdesc[300]{Computing methodologies~Machine learning}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% KEYWORDS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\keywords{Video memorability, Long-term memory, Scene understanding, Measurement protocol, Global video features, Attributes}

@@

@@

%objectives of the present study
While participating in the expansion of research on VM, the contributions of this paper are threefold:
\begin{itemize}
	\item We propose a new protocol that measures very long-term memory performances (from weeks to years after memorization) to collect ground truth data of good quality and make the corresponding dataset available for the research community (section \ref{section_dataset_construction}).
	\item We assess and analyze the collected data and provide useful insights on the understanding of VM (section \ref{study_memorability}).
	\item We build a computational model based on machine learning techniques which allows to predict a VM score for a given video. For this purpose, we investigate the use of various types of audio and visual features, ranging from low-level characteristics to emotional and (highly) semantic features (section \ref{mem-pred}).
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Previous work}
In this section, we review previous work on IM and VM prediction, focusing on annotation protocols and modelling aspects.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Measurement of memorability}
%work in psychology
Long-term memory has been studied for over a century in psychology, from multiple perspectives, since the seminal experimental studies of Ebbinghaus \cite{ebbinghaus_1913_memory} until the more recent neuro-imaging studies \cite{blumenfeld_2007_prefrontal}.
This work provided researchers, interested in computational understanding of IM and VM, with several memory tests (see \cite{richardson_1988_measures} for an extensive overview), such as the recognition test \cite{isola_2011_makes,khosla_2015_understanding,han_2015_learning} or the textual questions-based recall survey \cite{shekhar_2017_show}.
It also demonstrated that humans have an extensive long-term visual memory, which enables them to recall a great amount of images \cite{standing_1973_learning} and image details \cite{brady_2008_visual}, as well as of videos \cite{furman_2007_they}.
This is one of the observations at the origin of the work on memorability in computer science \cite{isola_2011_makes}.
Several factors have also been highlighted for their critical influence on long-term memory, such as emotion \cite{kensinger_2008_memory}, semantics \cite{quillan_1966_semantic}, several demographic factors \cite{cohendet_2016_using}, memory re-evocation \cite{nadel_1997_memory}, or passage of time \cite{mcgaugh_2000_memory}.
These different factors are all important to better understand memorability, and are found valuable computational features for IM and VM prediction.

@@

%Shekhar et al.
In another earlier approach for VM measurement, the participants performed a crowdsourcing experiment that consisted of a free viewing task during which they saw a series of videos, followed by a recall task in which they had to answer textual questions (such as: "Did you see a man juggling?", "Did you see a car on road?") \cite{shekhar_2017_show}.
The major drawback of the study comes from the use of questions instead of a classic visual recognition task.
Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty.
The authors tested the complexity of the chosen textual questions using the Flesch-Kincaid Grade Level readability metric, which is designed to quantify how difficult to understand an English text is \cite{kincaid_1975_derivation}.
However, this measure might not be sufficient to represent the questions in all their complexities, such as the ease for a person to draw images from the words, to connect the words to the associated concepts in memory, to associate the words to the corresponding scene, etc.
Moreover, there is an unequal distribution of the reading comprehension of English among Amazon Mechanical Turk workers, who are not necessary native English speakers.
This problem of a potential unequal complexity of the question becomes even more important in view of the use by authors of the response time to calculate memorability scores, which might also critically depend on the complexity of the questions.
One will note that this potential bias could have also affected the measure of inter-human consistency.
Furthermore, the questions were handcrafted, and the choices for types of questions and videos were very limited (e.g., the question "Did you see a car on road?" implies that in a whole experimental session only one car on a road should appear to connect the memory performance to one particular video).
The authors also manually ensured that no textual questions nor videos in a session were similar in content.
This makes it difficult to generalize this protocol to the construction of an extensive dataset.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memorability score calculation}
%Each video has also been viewed as a filler by $10.5$ participants.
After collecting the data, we kept only the 660 videos that had been seen at least 4 times as targets (from the initial set of 700 videos).
On average, each video of our dataset had been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enter into the calculation of a memorability score.
We then assigned a memorability score to each video, defined as the correct recognition rate of the video when viewed as target.
The average percentage of correct detections for all participants was $46.71\%$ (stdev = 14.65\%), and the average false alarm rate (i.e., the percentage of answers on fillers) was $4.16\%$ (stdev = 5.27\%).
Figure \ref{fig:response_time_vs_memorability}(a) provides a distribution of the videos according to their degree of memorability.
The dataset is publicly released here\footnote{\url{https://www.technicolor.com/dream/research-innovation/movie-memorability-dataset}}. In addition to the videos and corresponding ground truth, we also provide the features used in section \ref{mem-pred}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Consistency analysis}
\label{ssec:human}
%global consistency calculation
We follow the method proposed in \cite{isola_2014_makes} to measure human consistency when assessing memorability of videos.
We randomly split our 104 participants into two independent groups of equal size, and calculate how well VM scores from the first group of participants match with VM scores from the second group.
Averaging over 25 random half-split trials, an average Spearman's rank correlation (i.e., a global human consistency) of $0.57$ is observed between these pairs of scores.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Neutral and typical videos}
In our experiment, participants were given clear instructions that they had to really recognize any video they were presented as already seen, and not only guess that a video was coming from a movie whose title was proposed in the questionnaire.
However, a number of participants reported difficulty in making the distinction between guessing and recognizing: sometimes, because they knew that a video belonged to a movie they had seen, they had the false feeling of recognizing it.
In this section, we perform an analysis to compare neutral videos, which contain no element that would enable participants to easily guess that a video belongs to a particular movie, and typical videos, that could be more or less easily related to the movie they belong to.

A Wilcoxon rank-sum test indicated that the memorability was lower for neutral ($Mdn=.20, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, with $Z=10.22, p<.00001$.
This result does not necessarily mean that participants tended to guess when it was possible to do so -- rather than simply recognize -- videos selected from movies they have seen.
Indeed, neutral videos contain far less contextual elements useful for recognition than typical ones, making the latter -- if more easily "guessable" -- at least more easily recognizable.
However, it should draw our attention to the possibility that our measure is not purely objective, but -- to some extent -- context dependent.
This would constitute the main weakness of our protocol, that one should counteract by adapted measures of control.

@@

In \cite{shekhar_2017_show}, the response time of the participants was taken to be the measure of VM.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{New manner to compute memorability scores: take into account time and FA}
%Une nouvelle manière pour prendre en compte les FA dans les scores de mémorabilité applicable à l'étude de la mémorabilité sur les images
%Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.
% The question here is: "Should we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the movie --> No' --> So the point here is just to provide an overview.

@@

\subsection{Modelling and evaluation metric}
\label{model-eval}
We use the features discussed in Section \ref{feat-extr} to train a Support Vector Regression (SVR) model for the VM score prediction. The choice of SVR is guided by the nature of the problem as well as by the small size of the dataset. We have chosen to go with the same regressor for all the features because our focus is mainly on identifying which features are more important for VM prediction. This way we ensure that the difference in performances is because of the features themselves. Note that, in addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating them into a single feature vector.
While performing such a concatenation, we use the low-dimension version of the features for C3D and SentiBank features, obtained after applying a dimension reduction method (PCA) to the original set. In the experiments where we use PCA for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions. Note that we investigated other models like LSTM, to model the sequential nature of videos. The results were not very encouraging and this could be because of the relatively smaller size of our dataset.

@@

As it can be seen, visual semantic features derived from image captioning clearly offer better prediction results compared to all other considered features. This is not surprising as they capture visual attributes along the scene, which are known to play an important role in human memory. This is also inline with a previous study in IM \cite{squalli_2017_deep} where image captioning-based features helped better predict IM than the conventional CNN features. %These best prediction results (i.e., 0.31 and 0.38 for the test sets) are quite close to the human consistency shown in Figure \ref{fig:human_consistency}b (i.e. about 0.36 when having 4 annotations per video and about 0.47 when having 8 annotations per video).
When predicting VM on the test set, C3D features showed to be quite effective to the task as they encode the visual spatio-temporal information of the video. On the contrary, audio information captured by AudioSet features does not seem to be enough for VM prediction. One of our initial hypotheses was that emotion would play an important role in VM, supported by literature from psychology \cite{emo-mem}.
However, when observing the results in Table \ref{res-4-10-ann}, both Affect \cite{affect} and SentiBank \cite{sb-feat} performed quite poorly compared to the image captioning features.
%Observing the first set of results in Table \ref{res-4-10-ann}, where we have at least 4 annotations per video, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set. 
%Image captions capture the visual semantics in the video, while C3D features encode the visual spatio-temporal information.
%These results are consistent with previous results on predicting image memorability \cite{squalli_2017_deep}.
%Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
%Other features like AudioSet come close to the performance of the C3D, but only audio signal does not seem to be enough for predicting VM.
Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features. One of the reasons for this could be that there is a lot of redundance when combining all the features into a single feature vector, or it could be that the size of the combined feature vector is too big when compared to the size of the dataset. Thus, in future work we could look at selectively combining the features to investigate if that improves the performance.
